#ENERGY METRICS
#Source energy: metric used for determining Star Score
##represents the total amount of raw fuel that is required to operate the building. takes all energy use into account
#Site energy: amount of heat and electricity consumed by a building as reflected in your utility bills
# Weather normalized energy: the energy your building would have used under average conditions 
#WATER METRICS


```{r}
getwd()
setwd("/Users/madelyncaufield/Desktop/Sustainability Project")
df.raw = read.csv(file = "Existing_Buildings_Energy___Water_Efficiency__EBEWE__Program.csv")
head(df.raw)
names(df.raw)
sum(is.na(df.raw)) #95296 na's. might want to remove. 
```

```{r}
dff = na.omit(df.raw) #omitting NAs
anyNA(dff)
dff

df.NoNA=dff[-which(dff$TOTAL.WATER.USE..kgal..=="Not Available"),]



```

```{r}
library(plyr)

dff[which(dff$PROGRAM.YEAR == 2016), ] 
df.2016 = dff[which(dff$PROGRAM.YEAR == 2016), ] 
df.2016
df = data.frame(df.2016)
df
df.new<-df[c("Source.EUI..kBtu.ft..", "LADBS.Building.Category")]
df.new
head(df.new)

library(mltools)
df.new[, "energy.source.2"] <- bin_data(df.new$Source.EUI..kBtu.ft.., bins=3, binType = "quantile")
df.new #bins with equal number observations in each group

df.new.order = order(df.new$energy.source.2) 
df.new[df.new.order,]
df = df.new[df.new.order,] 

count(df.NoNA$LADBS.Building.Category)
count(df.NoNA$Source.EUI..kBtu.ft..)

library(devtools)
library(easyGgplot2)
source.energy.num = as.numeric(df.new$TOTAL.WATER.USE..kgal.)
energyplot = ggplot2.barplot(data=df.new, xName = 'LADBS.Building.Category', yName = 'source.energy.num',
                orientation="horizontal")
print(energyplot + labs(y="Total Water Use (kgal) 2019", x = "LADBS Building Category 2019"))

susn101=df.new[which(df.new$LADBS.Building.Category=="100,000+ sqft"),]
susn102=df.new[which(df.new$LADBS.Building.Category=="100,000+ sqft (CITY OWNED BUILDINGS)"),]
susn103=df.new[which(df.new$LADBS.Building.Category=="15,000 to 19,999 sqft (CITY OWNED BUILDING)"),]
susn104=df.new[which(df.new$LADBS.Building.Category=="20,000 to 49,999 sqft"),]
susn105=df.new[which(df.new$LADBS.Building.Category=="20,000 to 49,999 sqft (CITY OWNED BUILDINGS)"),]
susn106=df.new[which(df.new$LADBS.Building.Category=="50,000 to 99,999 sqft"),]
susn107=df.new[which(df.new$LADBS.Building.Category=="50,000 to 99,999 sqft (CITY OWNED BUILDINGS)"),]
susn108=df.new[which(df.new$LADBS.Building.Category=="7,500 to 14,999 sqft (CITY OWNED BUILDING)"),]

susn101$Source.EUI..kBtu.ft..=as.numeric(susn101$TOTAL.WATER.USE..kgal.)
susn102$Source.EUI..kBtu.ft..=as.numeric(susn102$TOTAL.WATER.USE..kgal.)
susn103$Source.EUI..kBtu.ft..=as.numeric(susn103$TOTAL.WATER.USE..kgal.)
susn104$Source.EUI..kBtu.ft..=as.numeric(susn104$TOTAL.WATER.USE..kgal.)
susn105$Source.EUI..kBtu.ft..=as.numeric(susn105$TOTAL.WATER.USE..kgal.)
susn106$Source.EUI..kBtu.ft..=as.numeric(susn106$TOTAL.WATER.USE..kgal.)
susn107$Source.EUI..kBtu.ft..=as.numeric(susn107$TOTAL.WATER.USE..kgal.)
susn108$Source.EUI..kBtu.ft..=as.numeric(susn108$STOTAL.WATER.USE..kgal.)

par(mfrow=c(3,3))
hist(susn101$Source.EUI..kBtu.ft..,main="100,000+ sqft", breaks = 25, xlab = "Total Water Use (kgal) 2019")
hist(susn102$Source.EUI..kBtu.ft..,main="100,000+ sqft (CITY OWNED BUILDINGS)", breaks = 25, xlab = "Total Water Use (kgal) 2019")
hist(susn103$Source.EUI..kBtu.ft..,main="15,000 to 19,999 sqft (CITY OWNED BUILDING)", breaks = 25, xlab = "Total Water Use (kgal) 2019")
hist(susn104$Source.EUI..kBtu.ft..,main="20,000 to 49,999 sqft", breaks = 25, xlab = "Total Water Use (kgal) 2019")
hist(susn105$Source.EUI..kBtu.ft..,main="20,000 to 49,999 sqft (CITY OWNED BUILDINGS)", breaks = 25, xlab = "Total Water Use (kgal) 2019")
hist(susn106$Source.EUI..kBtu.ft..,main="50,000 to 99,999 sqft", breaks = 25, xlab = "Total Water Use (kgal) 2019")
hist(susn107$Source.EUI..kBtu.ft..,main="50,000 to 99,999 sqft (CITY OWNED BUILDINGS)", breaks = 25, xlab = "Total Water Use (kgal) 2019")
hist(susn108$Source.EUI..kBtu.ft..,main="7,500 to 14,999 sqft (CITY OWNED BUILDING", breaks = 25, xlab = "Total Water Use (kgal) 2019")
```

```{r}
df.2016.order = order(df.2016$Source.EUI..kBtu.ft..) 
df.2016[df.2016.order,]
df = df.2016[df.2016.order,] #source energy use intensity in order from least to greatest

library(psych)
library("ggplot2")
library(devtools)
library(corrplot)
library(lattice)
library(car)
library(pastecs)
require(qqplotr)
library(effects)
library(lattice)
library(foreign)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(fastICA)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(pan)
library(mgcv)
library(DAAG)
library(MASS)

p1 = powerTransform(Source.EUI..kBtu.ft.. ~ 1, data = df.2016, family = "bcPower")


transformed_energy = bcPower(with(df.2016, Source.EUI..kBtu.ft..),
                       coef(p1, round = TRUE))
transformed_energy
plot(transformed_LotArea)
```


```{r}
library(devtools)
library(easyGgplot2)

 
head(df.NoNA)
table(df.NoNA$PROPERTY.TYPE)
table(df.NoNA$Source.EUI..kBtu.ft..)

df = data.frame(df.NoNA)
df.new<-df[c("Source.EUI..kBtu.ft..", "PROPERTY.TYPE")]
df.new
head(df.new)
ggplot2.barplot(data=df.new, xName = 'PROPERTY.TYPE', yName = 'Source.EUI..kBtu.ft..',
                orientation="horizontal")

library(mltools)
df.new[, "energy.source.2"] <- bin_data(df.new$Source.EUI..kBtu.ft.., bins=3, binType = "quantile")
df.new #bins with equal number observations in each group

df.new.order = order(df.new$energy.source.2) 
df.new[df.new.order,]
df = df.new[df.new.order,] 

library("writexl")
write_xlsx(df,"/Users/madelyncaufield/Desktop/Sustainability Project\\prop_energy.xlsx")

```

```{r}
library(corrplot)
C02 = as.numeric(df.NoNA$CARBON.DIOXIDE.EMISSIONS..Metric.Ton.CO2e.)
Source.energy = as.numeric(df.NoNA$Source.EUI..kBtu.ft..)
corrplot(cor(df.NoNA[c('CO2', 'source.energy')]))
numeric_col <- sapply(df.NoNA, is.numeric)
CO2_numeric <- df.NoNA[, numeric_col]
Correlation <- cor(CO2_numeric, use="pairwise.complete.obs")
corrplot(Correlation)



df.cor = cor(df.new)
df.cor
corrplot(df.cor, tl.cex=0.8, tl.srt=20, tl.col="black")
```

```{r}
df.new %>%
  filter(ENERGY.STAR.SCORE > 74)
df.new[df.new$ENERGY.STAR.SCORE > 74, ]

library("writexl")
write_xlsx(df,"/Users/madelyncaufield/Desktop/Sustainability Project\\prop_energy.xlsx")

library("readxl")
read_excel("/Users/madelyncaufield/Desktop/Sustainability Project\\prop_energy.xlsx")
```